Function to find association b/w n significant eigen genes with Diagnosis, Diagnosis.Gender, ApoE

### 2017-09-20 code/Rmd/networkAssociationTest.Rmd
Added total effect size randomisation and sanity modifications

---
title: "Function to find association b/w n significant eigene genes with Diagnosis"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')

@@

# Load Libraries
library(CovariateAnalysis)
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

@@

### Data download
```{r download.data, cache=TRUE}
# Download data
MODULES = synTableQuery('SELECT * FROM syn10338156')@values

@@

COVARIATES = downloadFile('syn10160582')
COVARIATES$BrainRegion[COVARIATES$BrainRegion == 'CER'] = 'CBE'

@@

EXPR = downloadFile('syn10160584')
gene.id.map = data.frame(ID = EXPR$Gene.ID, Gene.ID = EXPR$Gene.ID) %>%
  tidyr::separate(ID, c('GeneID','position'), sep = '\\.')
EXPR = left_join(EXPR, gene.id.map)
rownames(EXPR) = EXPR$GeneID
EXPR$Gene.ID = NULL
EXPR$GeneID = NULL

MODULE.OVRLP = synTableQuery('SELECT * FROM syn10339153')@values

MODULE.ENRICH = synTableQuery('SELECT * FROM syn10492048')@values

MAGMA.ENRICH = synTableQuery('SELECT * FROM syn10380432')@values
```

```{r fxn}
# Function to find association statistics
findAssociationStatistics <- function(x, COVARIATES, EXPR, FORMULA){

  covar = dplyr::filter(COVARIATES, BrainRegion == unique(x$brainRegion))
  expr = EXPR[x$GeneID,covar$SampleID]

  result = list( assoc.stat = data.frame(total.eff.sz = 0,
                                         ngenes = length(unique(x$GeneID)),
                                         npc = 0,
                                         npc.sig = 0,
                                         pc1.eff.sz = 0,
                                         pc1.fdr = 1),
                 anova.stat = data.frame() )

@@

  pc = prcomp(expr, center = T, scale. = T)
  pc.sdev = pc$sdev^2/sum(pc$sdev^2)
  ind = which(pc.sdev >= 0.01)
  pc.sdev = pc.sdev[ind]
  pc.rot = pc$rotation[,ind, drop = F]

  getAnovaAssociationStats <- function(dat, formula){
    mdl = as.data.frame(anova(lm(formula = formula, data = dat)))
    mdl$fdr = p.adjust(mdl[,'Pr(>F)'])
    eff.sz = mdl['Diagnosis','Sum Sq']/sum(mdl[,'Sum Sq'], na.rm = T)
    fdr = mdl['Diagnosis','fdr']
    return(data.frame(eff.sz = eff.sz, fdr = fdr))
  }

  anova.stat = apply(pc.rot, 2, function(x, covar, formula){ 
    dat = cbind(data.frame(PC = x), covar)
    getAnovaAssociationStats(dat, formula)
  }, covar, FORMULA) %>%
    rbindlist(use.names = T, fill = T)
  anova.stat$fve = pc.sdev

@@

  result = data.frame(total.eff.sz = sum(total.eff.sz$ws, na.rm = T),
                      ngenes = length(unique(x$GeneID)),
                      npc = dim(anova.stat)[1],
                      npc.sig = sum(anova.stat$fdr <= 0.05),
                      pc1.eff.sz = anova.stat$eff.sz[1],
                      pc1.fdr = anova.stat$fdr[1])

@@

### Associaiton with Diagnosis
```{r Dx, fig.height=7, fig.width=15}
# Find assocition b/w Diagnosis category with n eigen genes
FORMULA = as.formula('PC~Diagnosis+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED+Gender')
assoc.stat.diagnosis = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

Dx = lapply(assoc.stat.diagnosis, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique())
Dx$pc1.fdr.l = Dx$pc1.fdr <= 0.05

@@

p = ggplot(Dx, aes(x = method, y = total.eff.sz, color = method)) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx, aes(x = method, y = pc1.eff.sz, color = -log10(pc1.fdr))) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('PC1 Effect Size')
p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')
```
Top 5 modules per method
```{r Dx.1, fig.height=15, fig.width=15}
# Top 5 in each method
h = lapply(unique(Dx$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion, method) %>%
    dplyr::top_n(5, total.eff.sz) %>%
    dplyr::top_n(5, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)

  getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
}, Dx, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```
Top 5 modules any method
```{r Dx.2, fig.height=15, fig.width=15}
# Top 5 any method
h = lapply(unique(Dx$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion) %>%
    dplyr::top_n(25, total.eff.sz) %>%
    dplyr::top_n(25, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)
  print(summary(factor(imp.mods$method)))

@@

### Associaiton with Diagnosis.Gender
```{r Dx.Gender, fig.height=7, fig.width=15}
# Find assocition b/w Diagnosis.Gender category with n eigen genes
COVARIATES$Diagnosis.Gender = paste(COVARIATES$Diagnosis, COVARIATES$Gender, sep= '.') %>%
  as.factor
FORMULA = as.formula('PC~Diagnosis.Gender+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED')
assoc.stat.diagnosis.gender = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

Dx.gender = lapply(assoc.stat.diagnosis.gender, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique())
Dx.gender$pc1.fdr.l = Dx.gender$pc1.fdr <= 0.05

@@

p = ggplot(Dx.gender, aes(x = method, y = total.eff.sz, color = method)) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx.gender, aes(x = method, y = pc1.eff.sz, color = -log10(pc1.fdr))) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('PC1 Effect Size')
p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')
```
Top 5 modules per method
```{r Dx.gender.1, fig.height=15, fig.width=15}
# Top 5 in each method
h = lapply(unique(Dx.gender$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion, method) %>%
    dplyr::top_n(5, total.eff.sz) %>%
    dplyr::top_n(5, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)

  getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
}, Dx.gender, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```
Top 5 modules any method
```{r Dx.gender.2, fig.height=15, fig.width=15}
# Top 5 any method
h = lapply(unique(Dx.gender$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion) %>%
    dplyr::top_n(25, total.eff.sz) %>%
    dplyr::top_n(25, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)
  print(summary(factor(imp.mods$method)))

@@

### Associaiton with ApoE
```{r ApoE, fig.height=7, fig.width=15}
# Find assocition b/w ApoE category with n eigen genes
FORMULA = as.formula('PC~BrainRegion.ApoE+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED+Gender')
assoc.stat.apoe = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

ApoE = lapply(assoc.stat.apoe, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique())
ApoE$pc1.fdr.l = ApoE$pc1.fdr <= 0.05

@@

p = ggplot(ApoE, aes(x = method, y = total.eff.sz, color = method)) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

@@

# Combine and rank all the information
scr1 = Dx %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(rnk = rank(total.eff.sz),
                    rnk = rnk/max(rnk)) %>%
      dplyr::arrange(desc(rnk))%>%
      dplyr::group_by(ModuleNameFull) %>%
      dplyr::top_n(1, rnk) %>%
      dplyr::top_n(1, -log10(pc1.fdr))
  }) %>%
  dplyr::mutate(Dx = rank(rnk),
                Dx = Dx/max(Dx)) %>%
  dplyr::select(ModuleNameFull, brainRegion, method, ModuleName, Dx)

@@

ggplot(all.rank, aes(x = other, y = magma))+geom_point()+facet_grid(.~brainRegion)+geom_smooth(method = 'lm')
```
Top 5 modules per method
```{r imp.mdl.1, fig.height=15, fig.width=15}
# Top 5 in each method
h = lapply(unique(all.rank$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion, method) %>%
    dplyr::top_n(5, rnk) %>%
    dplyr::top_n(5, Dx) %>%
    dplyr::filter(brainRegion == br)

  getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
}, all.rank, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```
Top 5 modules any method
```{r imp.mdl.2, fig.height=15, fig.width=15}
# Top 5 any method
h = lapply(unique(all.rank$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion) %>%
    dplyr::top_n(25, rnk) %>%
    dplyr::top_n(25, Dx) %>%
    dplyr::filter(brainRegion == br)
  print(summary(factor(imp.mods$method)))

@@

### Modules missing enrichment scores
```{r missing}
writeLines('Modules missing enrich enrichment scores')
filter(Dx, ModuleNameFull %in% unique(setdiff(Dx$ModuleNameFull, MODULE.ENRICH$ModuleNameFull))) %>%
  kable

writeLines('Modules missing MAGMA enrichment scores')
filter(Dx, ModuleNameFull %in% unique(setdiff(Dx$ModuleNameFull, MAGMA.ENRICH$Set))) %>%
  kable
```

>>>>

---
title: "Function to find association b/w n significant eigen genes with Diagnosis, Diagnosis.Gender, ApoE"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')

@@

# Load Libraries
library(CovariateAnalysis) #devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

@@

### Data download
```{r download.data, cache=FALSE, include=FALSE}
# Download data
MODULES = synTableQuery('SELECT * FROM syn10338156')@values
all.used.ids = c('syn10338156')

@@

COVARIATES = downloadFile('syn10160582')
all.used.ids = c(all.used.ids, 'syn10160582')
COVARIATES$BrainRegion[COVARIATES$BrainRegion == 'CER'] = 'CBE'

@@

EXPR = downloadFile('syn10160584')
all.used.ids = c(all.used.ids, 'syn10160584')
gene.id.map = data.frame(ID = EXPR$Gene.ID, Gene.ID = EXPR$Gene.ID) %>%
  tidyr::separate(ID, c('GeneID','position'), sep = '\\.')
EXPR = left_join(EXPR, gene.id.map)
rownames(EXPR) = EXPR$GeneID
EXPR$Gene.ID = NULL
EXPR$GeneID = NULL

MODULE.OVRLP = synTableQuery('SELECT * FROM syn10339153')@values
all.used.ids = c(all.used.ids, 'syn10339153')
MODULE.ENRICH = synTableQuery('SELECT * FROM syn10492048')@values
all.used.ids = c(all.used.ids, 'syn10492048')

MAGMA.ENRICH = synTableQuery('SELECT * FROM syn10380432')@values
all.used.ids = c(all.used.ids, 'syn10380432')
```

```{r fxn}
# Function to find association statistics
findAssociationStatistics <- function(x, covars, exprs, frmla){

  covar = dplyr::filter(covars, BrainRegion == unique(x$brainRegion))
  expr = exprs[x$GeneID,covar$SampleID]

  result = list( assoc.stat = data.frame(total.eff.sz = 0,
                                         total.pval = 1,
                                         ngenes = length(unique(x$GeneID)),
                                         npc = 0,
                                         npc.sig = 0,
                                         pc1.eff.sz = 0,
                                         pc1.fdr = 1),
                 anova.stat = data.frame() )

@@

  # Function to get anova association stats
  getAnovaAssociationStats <- function(dat, frmla){
    mdl = as.data.frame(anova(lm(formula = frmla, data = dat)))
    mdl$fdr = p.adjust(mdl[,'Pr(>F)'])
    eff.sz = mdl['Diagnosis','Sum Sq']/sum(mdl[,'Sum Sq'], na.rm = T)
    fdr = mdl['Diagnosis','fdr']
    return(data.frame(eff.sz = eff.sz, fdr = fdr))
  }

  # Function to get PCs
  getPCs <- function(expr){
    pc = prcomp(expr, center = T, scale. = T)
    pc.sdev = pc$sdev^2/sum(pc$sdev^2)
    ind = which(pc.sdev >= 0.01)
    pc.sdev = pc.sdev[ind]
    pc.rot = pc$rotation[,ind, drop = F]

    return(list(sdev = pc.sdev, rot = pc.rot))
  }

  # Function to get PCs
  bootstrapAssociationStats <- function(expr, covar, frmla){
    pc = getPCs(expr)

    anova.stat = apply(pc$rot, 2, function(x, covar, frmla){ 
      dat = cbind(data.frame(PC = x), covar)
      getAnovaAssociationStats(dat, frmla)
    }, covar, frmla) %>%
      rbindlist(use.names = T, fill = T)
    anova.stat$fve = pc$sdev

@@

    sum(total.eff.sz$ws, na.rm = T)
  }

  pc = getPCs(expr)

  anova.stat = apply(pc$rot, 2, function(x, covar, frmla){ 
    dat = cbind(data.frame(PC = x), covar)
    getAnovaAssociationStats(dat, frmla)
  }, covar, frmla) %>%
    rbindlist(use.names = T, fill = T)
  anova.stat$fve = pc$sdev

  total.eff.sz = dplyr::filter(anova.stat, fdr <= 0.05) %>%
    dplyr::mutate(ws = eff.sz * fve) %>%
    dplyr::select(ws) 
  total.eff.sz = sum(total.eff.sz$ws, na.rm = T)

  rn.eff.sz = plyr::laply(1:100, .fun = function(i, expr, l, covar, frmla){
    tryCatch({ 
      bootstrapAssociationStats(expr[sample(1:dim(expr)[1], l), ], covar, frmla) 
    }, error = function(e){ 0 })
  }, exprs[,covar$SampleID], length(x$GeneID), covar, frmla, .parallel = T)

  result = data.frame(total.eff.sz = total.eff.sz,
                      total.pval = sum(rn.eff.sz >= total.eff.sz, na.rm = T)/length(rn.eff.sz),
                      ngenes = length(unique(x$GeneID)),
                      npc = dim(anova.stat)[1],
                      npc.sig = sum(anova.stat$fdr <= 0.05),
                      pc1.eff.sz = anova.stat$eff.sz[1],
                      pc1.fdr = anova.stat$fdr[1])

@@

### Associaiton with Diagnosis
```{r Dx, fig.height=5, fig.width=10}
# Find assocition b/w Diagnosis category with n eigen genes
FORMULA = as.formula('PC~Diagnosis+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED+Gender')
assoc.stat.diagnosis = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

Dx = lapply(assoc.stat.diagnosis, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique()) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(total.fdr = p.adjust(total.pval))
  })

@@

p = ggplot(Dx, aes(x = method, y = total.eff.sz, color = total.fdr)) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('Total Effect Size')
p = p + scale_color_continuous(low = 'red', high = 'white')
p + theme(legend.position = 'top')

p = ggplot(Dx, aes(x = method, y = pc1.eff.sz, color = -log10(pc1.fdr))) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('PC1 Effect Size')
p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')
```
Top 5 modules per method
```{r Dx.1, fig.height=12, fig.width=12}
# Top 5 in each method
h = lapply(unique(Dx$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion, method) %>%
    dplyr::filter(total.fdr <= 0.05) %>%
    dplyr::top_n(5, total.eff.sz) %>%
    dplyr::top_n(5, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)

  getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
}, Dx, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```
Top 5 modules any method
```{r Dx.2, fig.height=12, fig.width=12}
# Top 5 any method
h = lapply(unique(Dx$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion) %>%
    dplyr::filter(total.fdr <= 0.05) %>%
    dplyr::top_n(25, total.eff.sz) %>%
    dplyr::top_n(25, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)
  print(summary(factor(imp.mods$method)))

@@

### Associaiton with Diagnosis.Gender
```{r Dx.Gender, fig.height=5, fig.width=10}
# Find assocition b/w Diagnosis.Gender category with n eigen genes
COVARIATES$Diagnosis.Gender = paste(COVARIATES$Diagnosis, COVARIATES$Gender, sep= '.') %>%
  as.factor
FORMULA = as.formula('PC~Diagnosis.Gender+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED')
assoc.stat.diagnosis.gender = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

Dx.gender = lapply(assoc.stat.diagnosis.gender, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique()) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(total.fdr = p.adjust(total.pval))
  })

@@

p = ggplot(Dx.gender, aes(x = method, y = total.eff.sz, color = total.fdr)) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('Total Effect Size')
p = p + scale_color_continuous(low = 'red', high = 'white')
p + theme(legend.position = 'top')

p = ggplot(Dx.gender, aes(x = method, y = pc1.eff.sz, color = -log10(pc1.fdr))) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('PC1 Effect Size')
p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')
```
Top 5 modules per method
```{r Dx.gender.1, fig.height=12, fig.width=12}
# Top 5 in each method
h = lapply(unique(Dx.gender$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion, method) %>%
    dplyr::filter(total.fdr <= 0.05) %>%
    dplyr::top_n(5, total.eff.sz) %>%
    dplyr::top_n(5, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)

  getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
}, Dx.gender, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```
Top 5 modules any method
```{r Dx.gender.2, fig.height=12, fig.width=12}
# Top 5 any method
h = lapply(unique(Dx.gender$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion) %>%
    dplyr::filter(total.fdr <= 0.05) %>%
    dplyr::top_n(25, total.eff.sz) %>%
    dplyr::top_n(25, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)
  print(summary(factor(imp.mods$method)))

@@

### Associaiton with ApoE
```{r ApoE, fig.height=5, fig.width=10}
# Find assocition b/w ApoE category with n eigen genes
FORMULA = as.formula('PC~BrainRegion.ApoE+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED+Gender')
assoc.stat.apoe = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

ApoE = lapply(assoc.stat.apoe, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique()) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(total.fdr = p.adjust(total.pval))
  })

@@

p = ggplot(ApoE, aes(x = method, y = total.eff.sz, color = total.fdr)) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('Total Effect Size')
p = p + scale_color_continuous(low = 'red', high = 'white')
p + theme(legend.position = 'top')

@@

# Combine and rank all the information
scr1 = Dx %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(rnk = rank(rank(total.eff.sz) + rank(-log10(total.fdr))),
                    rnk = rnk/max(rnk)) %>%
      dplyr::arrange(desc(rnk))%>%
      dplyr::group_by(ModuleNameFull) %>%
      dplyr::top_n(1, rnk) %>%
      dplyr::top_n(1, -log10(pc1.fdr))
  }) %>%
  dplyr::mutate(Dx = rank(rnk),
                Dx = Dx/max(Dx)) %>%
  dplyr::select(ModuleNameFull, brainRegion, method, ModuleName, Dx)

@@

ggplot(all.rank, aes(x = other, y = magma))+geom_point()+facet_grid(.~brainRegion)+geom_smooth(method = 'lm')
```
Top 5 modules per method
```{r imp.mdl.1, fig.height=12, fig.width=12}
# Top 5 in each method
h = lapply(unique(all.rank$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion, method) %>%
    dplyr::top_n(5, rnk) %>%
    dplyr::top_n(5, Dx) %>%
    dplyr::filter(brainRegion == br)

  getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
}, all.rank, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```
Top 5 modules any method
```{r imp.mdl.2, fig.height=12, fig.width=12}
# Top 5 any method
h = lapply(unique(all.rank$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion) %>%
    dplyr::top_n(25, rnk) %>%
    dplyr::top_n(25, Dx) %>%
    dplyr::filter(brainRegion == br)
  print(summary(factor(imp.mods$method)))

@@

### Modules missing enrichment scores
```{r missing, include=FALSE}
writeLines('Modules missing enrich enrichment scores')
filter(Dx, ModuleNameFull %in% unique(setdiff(Dx$ModuleNameFull, MODULE.ENRICH$ModuleNameFull))) %>%
  dplyr::select(ModuleNameFull) %>% unique() %>% unlist() %>% paste(collapse = ', ') %>% print()

writeLines('Modules missing MAGMA enrichment scores')
filter(Dx, ModuleNameFull %in% unique(setdiff(Dx$ModuleNameFull, MAGMA.ENRICH$Set))) %>%
  dplyr::select(ModuleNameFull) %>% unique() %>% unlist() %>% paste(collapse = ', ') %>% print()
```

### Store results in synapse
```{r syn.store, cache = FALSE, include=FALSE}
CODE = Folder('Network Association Tests', parentId = "syn10306449")
CODE = synStore(CODE)

write.table(Dx, file = 'AssociaitonStatWithDx.tsv', sep = '\t', quote = F, row.names = F)
OBJ = File('AssociaitonStatWithDx.tsv', name = 'Association with Dx', parentId = CODE@properties$id)
OBJ = synStore(OBJ, used = all.used.ids, executed = thisFile, activityName = activityName)

write.table(Dx.gender, file = 'AssociaitonStatWithDxGender.tsv', sep = '\t', quote = F, row.names = F)
OBJ1 = File('AssociaitonStatWithDxGender.tsv', name = 'Association with Dx.Gender', parentId = CODE@properties$id)
OBJ1 = synStore(OBJ1, used = all.used.ids, executed = thisFile, activityName = activityName)

write.table(ApoE, file = 'AssociaitonStatWithApoE4.tsv', sep = '\t', quote = F, row.names = F)
OBJ2 = File('AssociaitonStatWithApoE4.tsv', name = 'Association with ApoE4', parentId = CODE@properties$id)
OBJ2 = synStore(OBJ2, used = all.used.ids, executed = thisFile, activityName = activityName)

save(list = c('assoc.stat.diagnosis', 'assoc.stat.diagnosis.gender', 'assoc.stat.apoe'), file = 'AllAssociationResults.RData')
OBJ3 = File('AllAssociationResults.RData', name = 'All Assocition Statistics (RData)', parentId = CODE@properties$id)
OBJ3 = synStore(OBJ3, used = all.used.ids, executed = thisFile, activityName = activityName)
```
